---
title: "Metadata on available specificity matrices/dataframes"
author: "Regina H. Reynolds"
output: 
  html_document:
    theme: paper
    highlight: kate
    df_print: paged
    toc: true
    toc_float: true
    number_sections: true
---

```{r setup, include=FALSE}

library(DT)
library(readxl)
library(EWCE)
library(ggplot2)
library(ggpubr)
library(tidyverse)
library(stringr)

```


# Descriptions for metadata columns
```{r metadata columns, echo=FALSE, out.width='100%', fig.pos='H'}

readxl::read_excel(path = "/home/rreynolds/projects/MarkerGenes/metadata/dataset_metadata.xlsx", sheet = "ColumnDescriptions") %>% 
  DT::datatable(rownames = FALSE,
                options = list(scrollX = TRUE),
                class = 'white-space: nowrap')

```

# Dataset metadata
```{r metadata, echo=FALSE, out.width='100%', fig.pos='H'}

readxl::read_excel(path = "/home/rreynolds/projects/MarkerGenes/metadata/dataset_metadata.xlsx", sheet = "Data") %>% 
  arrange(desc(data_type), resource_name) %>% 
  DT::datatable(rownames = FALSE,
                options = list(scrollX = TRUE),
                class = 'white-space: nowrap')

```

# Notes
- The majority of specificity matrices have been downloaded from external sources (primarily [Skene 2018](https://pubmed.ncbi.nlm.nih.gov/29785013/); with specificity matrices made available in the associated [GitHub repo](https://github.com/NathanSkene/MAGMA_Celltyping/tree/master/data)).
- A few specificity matrices have been specifically generated within the `MarkerGene` package, including:
    - Zeisel et al. 2018 (see: `r  xfun::embed_file(path = "/home/rreynolds/projects/LDSC_Regression/R/EWCE/Zeisel2018_CNSClusters_FormattingDataForEWCE.R", name = "Zeisel2018_CNSClusters_FormattingDataForEWCE.R", text = "Zeisel2018_CNSClusters_FormattingDataForEWCE.R")`)
    - Allen Brain Atlas 2018 **(only MTG)**. 
        - Downloaded from [AIBS](https://portal.brain-map.org/atlases-and-data/rnaseq/human-mtg-smart-seq).
        - Used gene-level (exonic) read count values.
        - For code on how this was generated, please refer to: [AIBS2018_GeneratingSImatrix.R](../data-raw/AIBS2018_GeneratingSImatrix.R). 
    - [Agarwal 2020](https://www.nature.com/articles/s41467-020-17876-0).
        - Data downloaded from [GEO](https://www.ncbi.nlm.nih.gov/geo/query/acc.cgi?acc=GSE140231). 
        - For code on how this was generated, please refer to: [Agarwal2020_GeneratingSImatrix.R](../data-raw/Agarwal2020_GeneratingSImatrix.R.R)
- Finally, a specificity dataframe was generated within the `MarkerGene` package for GTEx v8. 
    - This was generated by modifying code from the article, [Bryois 2020](https://pubmed.ncbi.nlm.nih.gov/32341526/), which was made available here: https://github.com/jbryois/scRNA_disease/blob/master/Code_Paper/Code_GTEx/get_GTEx_input.md. 
    - The primary modification was to include all genes, including non-protein-coding genes, which were previously excluded.
    - For all details on the generation of this specificity dataframe, please refer to: [GTEx_v8_specificity.html](../data-raw/GTEx_v8_specificity.html).
 
