---
title: "Marker lists"
author: "Regina H. Reynolds"
date: "6/14/2019"
output: 
  html_document:
    theme: paper
    highlight: kate
    df_print: paged
    toc: true
    toc_float: true
    number_sections: true
---

```{r setup, include=FALSE}

library(DT)
library(readxl)
library(EWCE)
library(ggplot2)
library(ggpubr)
library(tidyverse)
library(stringr)

```


# Descriptions for metadata columns
```{r metadata columns, echo=FALSE, out.width='100%', fig.pos='H'}

readxl::read_excel(path = "/home/rreynolds/projects/MarkerGenes/metadata/dataset_metadata.xlsx", sheet = "ColumnDescriptions") %>% 
  DT::datatable(rownames = FALSE,
                options = list(scrollX = TRUE),
                class = 'white-space: nowrap')

```

# Dataset metadata
```{r metadata, echo=FALSE, out.width='100%', fig.pos='H'}

readxl::read_excel(path = "/home/rreynolds/projects/MarkerGenes/metadata/dataset_metadata.xlsx", sheet = "Data") %>% 
  arrange(desc(list_or_SI_matrix), resource_name) %>% 
  DT::datatable(rownames = FALSE,
                options = list(scrollX = TRUE),
                class = 'white-space: nowrap')

```

- The majority of specificity matrices were downloaded from Skene 2018, but also generated two matrices from:
    - Zeisel et al. 2018 (see: `/home/rreynolds/projects/LDSC_Regression/R/EWCE/Zeisel2018_CNSClusters_FormattingDataForEWCE.R`)
    - Allen Brain Atlas 2019 (see: `/home/rreynolds/projects/MarkerGenes/misc_scripts/20190501_AIBS2019_GeneratingSImatrix.R`). Used gene-level (exonic) read count values. While intronic values were available, reckoned that these would tag pre-mRNA.
