---
title: "Dataset metadata"
author: 
  - name: Regina H. Reynolds
    affiliation:
    - UCL
    email: regina.reynolds.16@ucl.ac.uk
output: 
  BiocStyle::html_document:
    self_contained: yes
    toc: true
    toc_float: true
    toc_depth: 2
    code_folding: show
date: "`r doc_date()`"
package: "`r pkg_ver('MarkerGenes')`"
vignette: >
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}  
---

```{r setup, include=FALSE}

library(DT)
library(here)
library(readxl)
library(tidyverse)

```

# Descriptions for metadata columns

```{r metadata columns, echo=FALSE, out.width='100%', fig.pos='H'}

readxl::read_excel(
  path = here::here("metadata", "dataset_metadata.xlsx"), 
  sheet = "ColumnDescriptions"
  ) %>% 
  DT::datatable(rownames = FALSE,
                options = list(scrollX = TRUE),
                class = 'white-space: nowrap')

```

# Dataset metadata

```{r metadata, echo=FALSE, out.width='100%', fig.pos='H'}

readxl::read_excel(
  path = here::here("metadata", "dataset_metadata.xlsx"), 
  sheet = "Data"
  ) %>% 
  dplyr::arrange(desc(data_type), resource_name) %>% 
  DT::datatable(rownames = FALSE,
                options = list(scrollX = TRUE),
                class = 'white-space: nowrap')

```

# Generation of specificity matrices/dataframes

- The majority of specificity matrices have been downloaded from external sources (primarily [Skene 2018](https://pubmed.ncbi.nlm.nih.gov/29785013/); with specificity matrices made available in the associated [GitHub repo](https://github.com/NathanSkene/MAGMA_Celltyping/tree/master/data)).
- A few specificity matrices have been specifically generated within the `MarkerGene` package. These can be divided into their compatibility with EWCE versions:
    - Compatible with the defunct [EWCE v1.3.0](https://bioconductor.riken.jp/packages/3.5/bioc/html/EWCE.html) available on Bioconductor v3.5
        - Zeisel et al. 2018 (see: [Zeisel2018_CNSClusters_FormattingDataForEWCE.R](https://github.com/RHReynolds/MarkerGenes/blob/master/data-raw/Zeisel2018_CNSClusters_FormattingDataForEWCE.R))
        - Allen Brain Atlas 2018 **(only MTG)**. 
            - Downloaded from [AIBS](https://portal.brain-map.org/atlases-and-data/rnaseq/human-mtg-smart-seq).
            - Used gene-level (exonic) read count values.
            - For code on how this was generated, please refer to: [AIBS2018_GeneratingSImatrix.R](https://github.com/RHReynolds/MarkerGenes/blob/master/data-raw/AIBS2018_GeneratingSImatrix.R). 
        - [Agarwal 2020](https://www.nature.com/articles/s41467-020-17876-0).
            - Data downloaded from [GEO](https://www.ncbi.nlm.nih.gov/geo/query/acc.cgi?acc=GSE140231). 
            - For code on how this was generated, please refer to: [Agarwal2020_GeneratingSImatrix.R](https://github.com/RHReynolds/MarkerGenes/blob/master/data-raw/Agarwal2020_GeneratingSImatrix.R)
    - Compatible with the new [EWCE](https://nathanskene.github.io/EWCE/index.html) that is available in Bioconductor>=3.14
        - Allen Brain Atlas 2018 **(multiple cortical areas)**
            - Downloaded from [AIBS](https://portal.brain-map.org/atlases-and-data/rnaseq/human-multiple-cortical-areas-smart-seq).
            - Used gene-level (exonic) read count values.
            - For code on how this was generated, please refer to: [AIBS2019_GeneratingSImatrix.R](https://github.com/RHReynolds/MarkerGenes/blob/master/data-raw/AIBS2019_GeneratingSImatrix.R) and [AIBS2019_ctd_generation_description.html](https://github.com/RHReynolds/MarkerGenes/blob/master/data-raw/AIBS2019_ctd_generation_description.html)
- Finally, a specificity dataframe was generated within the `MarkerGene` package for GTEx v8. 
    - This was generated by modifying code from the article, [Bryois 2020](https://pubmed.ncbi.nlm.nih.gov/32341526/), which was made available here: https://github.com/jbryois/scRNA_disease/blob/master/Code_Paper/Code_GTEx/get_GTEx_input.md. 
    - The primary modification was to include all genes, including non-protein-coding genes, which were previously excluded.
    - For all details on the generation of this specificity dataframe, please refer to: [GTEx_v8_specificity.html](https://github.com/RHReynolds/MarkerGenes/blob/master/data-raw/GTEx_v8_specificity.Rmd).
 
